# brandr <a href = "https://danielvartan.github.io/brandr/"><img src = "man/figures/logo.png" align="right" width="120" /></a>

<!-- quarto render -->

```{r}
#| label: setup
#| include: false

library(badger)
library(brandr)
# library(rutils) # github.com/danielvartan/rutils

source(here::here("R", ".setup.R"))
```

<!-- badges: start -->
```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_repostatus("active"),
  badger::badge_cran_release(),
  badger::badge_doi(
    doi = "10.32614/CRAN.package.brandr",
    color = "1284C5"
  ),
  badger::badge_cran_download(type = "grand-total"),
  badger::badge_github_actions(action = "R-CMD-check.yaml"),
  badger::badge_codecov() |> rutils::shush(),
  # badger::badge_custom(
  #   "fair--software.eu",
  #   "%E2%97%8F%20%20%E2%97%8F%20%20%E2%97%8B%20%20%E2%97%8F%20%20%E2%97%8B-",
  #   "orange",
  #   "https://fair-software.eu"
  # ),
  # paste0(
  #   "[![FAIR checklist badge](https://fairsoftwarechecklist.net/badge.svg)]",
  #   "(https://fairsoftwarechecklist.net/v0.2?f=21&a=32113&i=32001&r=123)"
  # ),
  # badger::badge_custom( # CII Best Practices
  #   "openssf%20best%20practices",
  #   "Placeholder",
  #   "orange",
  #   "https://bestpractices.coreinfrastructure.org/"
  # ),
  badger::badge_license(
    "MIT",
    "green",
    url = "https://choosealicense.com/licenses/mit/"
  ),
  badger::badge_custom( # Contributor Covenant
    "Contributor%20Covenant",
    "2.1",
    "4baaaa",
    "CODE_OF_CONDUCT.md"
  )
)
```
<!-- badges: end -->

## Overview

`brandr` is an R package designed to facilitate brand identity management using the [brand.yml](https://posit-dev.github.io/brand-yml/) standard. It provides functions to consistently access and apply brand colors, typography, and other visual elements across your R projects.

Unlike [Posit](https://posit.co/)'s [`thematic`](https://rstudio.github.io/thematic/) package, `brandr` offers more refined control over brand assets.

> If you appreciate the project, consider giving it a star! (â­)

**Please note that `brandr` is not affiliated with [Posit](https://posit.co/) or the developers of [brand.yml](https://posit-dev.github.io/brand-yml/).**

## Installation

You can install the released version of `brandr` from [CRAN](https://CRAN.R-project.org/package=brandr) with:

```r
install.packages("brandr")
```

And the development version from [GitHub](https://github.com/) with:

```r
# install.packages("remotes")
remotes::install_github("danielvartan/brandr")
```

## Usage

`brandr` is equipped with several functions to help you use your `_brand.yml` file, like:

- [`color_brand_sequential()`](https://danielvartan.github.io/brandr/reference/color_brand_sequential.html), `color_brand_diverging()`, `color_brand_qualitative()`: Sequential, diverging, and qualitative brand color palettes
- [`scale_brand()`](https://danielvartan.github.io/brandr/reference/scale_brand.html), `scale_color_brand_d()`, `scale_color_brand_c()`, `scale_color_brand_b()`, `scale_fill_brand_d()`, `scale_fill_brand_c()`, `scale_fill_brand_b()`: Discrete, continuous, and binned brand color scales for [`ggplot2`](https://ggplot2.tidyverse.org/)
- [`get_brand_color()`](https://danielvartan.github.io/brandr/reference/get_brand_color.html): Get a brand color by name (e.g., primary)
- [`get_brand_color_tint()`](https://danielvartan.github.io/brandr/reference/get_brand_color_tint.html): Get a brand color tint
- [`get_brand_color_mix()`](https://danielvartan.github.io/brandr/reference/get_brand_color_mix.html): Get a mix of brand colors
- [`get_brand_font()`](https://danielvartan.github.io/brandr/reference/get_brand_font.html): Get a brand font by name (e.g., headings)

Example:

```{r}
#| label: readme-figure
#| warning: false

library(brandr)
library(ggplot2)
library(palmerpenguins)

penguins |>
  ggplot(aes(x = flipper_length_mm, fill = species)) +
    geom_histogram(alpha = 0.5, position = "identity") +
    scale_fill_brand_d() +
    labs(
      x = "Flipper Length (mm)",
      y = "Frequency",
      fill = "Species"
    )
```

Click [here](https://danielvartan.github.io/brandr/) to see the full list.

## Configuration

### Path to `_brand.yml`

`brandr` will always look for a `_brand.yml` file in the root directory of your project. If the file is not found, an error message will be displayed. You can also set the path to the file manually using the `options()` function:

```r
options(BRANDR_BRAND_YML = "PATH_TO_BRAND.YML")
```

### Brand Color Scales

To control the colors of each brand color scale, assign [hexadecimal](https://en.wikipedia.org/wiki/Web_colors) color codes in a `character` vector to the following options:

- `BRANDR_COLOR_SEQUENTIAL`: For sequential color scales
- `BRANDR_COLOR_DIVERGING`:  For diverging color scales
- `BRANDR_COLOR_QUALITATIVE`: For qualitative color scales

You can use `get_brand_color()` to get the hexadecimal color codes from the `_brand.yml` file.

Example:

```r
options(
 BRANDR_COLOR_SEQUENTIAL =
   get_brand_color(c("primary", "secondary")),
 BRANDR_COLOR_DIVERGING =
   get_brand_color(c("primary", "white", "secondary")),
 BRANDR_COLOR_QUALITATIVE =
   get_brand_color(c("primary", "secondary", "tertiary"))
)
```

## Setting Brand Fonts/Typefaces

Currently, `brandr` does not support the use of custom fonts directly. This feature is planned for future releases.

In the meantime, you can use Yixuan Qiu's [`showtext`](https://doi.org/10.32614/CRAN.package.showtext) R package to apply custom fonts/typefaces to your plots.

Below is an example of how to use the Google Font [Open Sans](https://fonts.google.com/specimen/Open+Sans) in all of your plots. This example assumes that the font files are located in a `ttf` folder within your project directory.

```r
library(brandr)
library(ggplot2)
library(here)
library(sysfonts)
library(showtext)

font_paths(here("ttf"))

font_add(
  family = "open-sans",
  regular = here("ttf", "opensans-regular.ttf"),
  bold = here("ttf", "opensans-bold.ttf"),
  italic = here("ttf", "opensans-italic.ttf"),
  bolditalic = here("ttf", "opensans-bolditalic.ttf"),
  symbol = NULL
)

showtext_auto()

theme_set(
  theme(
    text = element_text(
      color = get_brand_color("black"),
      family = "open-sans",
      face = "plain"
    )
  )
)
```

## License

```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_license(
    "MIT",
    "green",
    url = "https://choosealicense.com/licenses/mit/"
  )
)
```

`brandr` code is released under the [MIT license](https://opensource.org/license/mit).

## Contributing

```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_custom( # Contributor Covenant
    "Contributor%20Covenant",
    "2.1",
    "4baaaa",
    "CODE_OF_CONDUCT.md"
  )
)
```

Contributions are welcome, including bug reports. Take a moment to review the [Guidelines for Contributing](https://danielvartan.github.io/brandr/CONTRIBUTING.html).

[![](https://img.shields.io/static/v1?label=Sponsor&message=%E2%9D%A4&logo=GitHub&color=%23fe8e86)](https://github.com/sponsors/danielvartan)

You can also support the development of `brandr` by becoming a sponsor. Click [here](https://github.com/sponsors/danielvartan) to make a donation. Please mention `brandr` in your donation message.

## Acknowledgments

`brandr` brand identity is based on [brand.yml](https://posit-dev.github.io/brand-yml/) brand identity.
